---
title: Simulation Report
subtitle: |
    ROADMAP: RandOmised Arthroplasty infection worlDwide Multidomain Adaptive Platform trial - simulation report
description: |
    Investigator initiated, Randomised Embedded Multifactorial Adaptive Platform (REMAP) trial, conducted across multiple hospitals in several regions of the world.
date: last-modified
date-format: "D MMMM YYYY"
author: 
  - name: Mark Jones
    id: mj
    email: mark.jones1@sydney.edu.au
  - name: James Totterdell
    id: jt
    email: james.totterdell@sydney.edu.au
version: 0.1
sponsor: University of Newcastle, NSW, Australia
protocol-number: Version 1.1  01AUG2024
registration: todo
hrec: todo
ci1: Professor Joshua Davis  
ci2: Professor Laurens Manning
editor: source
bibliography: etc/refs.bib
csl: etc/elsevier-harvard.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

```{r, echo = F}
#| echo: false

# uml digs
library(nomnoml)
library(ggplot2)
suppressPackageStartupMessages(library(data.table))
library(tibble)
library(gt)
library(kableExtra)
suppressPackageStartupMessages(library(qs))
library(git2r)
```

{{< pagebreak >}}

::: summary
|     |        |
|:----|:------------|
|Study title:  |  ROADMAP: RandOmised Arthroplasty infection worlDwide Multidomain Adaptive Platform trial |
|Intervention: |  Surgery type, backbone antibiotic duration, extended prophylaxis, antibiotic type |   
|Study design:  |   Randomised Embedded Multifactorial Adaptive Platform trial | 
|Sponsor:  |    University of Newcastle, NSW, Australia | 
Protocol: |  Version 1.1  01AUG2024 |
|Registration:  |    todo | 
|HREC:  |   todo | 
|Study date of first consent:  |   todo | 
|Principal coordinating investigators:  |   Professor Joshua Davis  and Professor Laurens Manning | 
:::

<!-- 
Note that the above relies on the pandoc extension implemented in the lua file 
in the etc directory. It additionally relies on the presence of a custom style
in word called study summary. It will currently only work for word (because I
cannot be bothered to implement it in anything else at the moment).
-->


{{< pagebreak >}}

# Version history {.unlisted .unnumbered}

| Version    |   Date     | Change    |   Reason     |
|:----|:------------|:----|:------------|
| 0.1 | 09/2024 | First version | N/A |
| 0.2 | 12/2024 | Review 1 | Edits based on initial comments from TS and JT |


{{< pagebreak >}}

# Repository status {.unlisted .unnumbered}

\footnotesize
```{r}
#| echo: false
repo <- repository(path = ".")
summary(repo)
```
\normalsize

{{< pagebreak >}}

# Preface {.unlisted .unnumbered}

This simulation report documents the current set of reference simulations for ROADMAP.
The simulation report is an operational document that will be updated, as necessary, over the course of the study.



